### 2019-09-17 Thesis/pmonads.tex
Random commit.

Equivalently, a lax action is given by a functor $\blank.\blank \from \X \times \C \to \C$ together with natural transformations
\begin{mathpar}
  m_{x,y,a}\from x.y.a \to (x\tensor y).a
  \and
  e_a \from a \to I.a
\end{mathpar}
such that the following diagrams commute.
\begin{mathpar}
  \begin{tikzcd}[column sep=36pt]
    x.y.z.a \arrow[r, "x.m_{y,z,a}"] \arrow[d, "m_{x,y,z.a}"']
      & x.(y\tensor z).a \arrow[r, "m_{x,y\tensor z,a}"]
        & (x \tensor (y \tensor z)) . a \\
    (x\tensor y).z.a \arrow[r, "m_{x\tensor y,z,a}"]
      & ((x\tensor y)\tensor z) . a \arrow[ur, "\assoc_{x,y,z}.a"']
        &
  \end{tikzcd}
  \and
  \begin{tikzcd}
    x.a \arrow[r, "e_{x.a}"] \arrow[dr, "\lunit_x.a"']
      & I.x.a \arrow[d, "m_{I,x,a}"] \\
    %
      & (I\tensor x).a
  \end{tikzcd}
  \and
  \begin{tikzcd}
    x.a \arrow[r, "x.e_a"] \arrow[dr, "\runit_x.a"']
      & x.I.a \arrow[d, "m_{x,I,a}"] \\
    %
      & (x\tensor I) . a
  \end{tikzcd}
\end{mathpar}
Since lax actions generalize monads, we shall follow \Mellies \cite{ParametricMonads} and refer to them as \emph{parametric monads}.

@@

\begin{definition}
  Given \Mellies morphisms $f$ from $a$ to $b$ and $g$ from $b$ to $c$ given by
  \begin{mathpar}
    \tilde{f} \from a \to x.b
    \and
    \tilde{g} \from b \to y.c\,,
  \end{mathpar}
  their \emph{\Mellies composition} is given by the composite
  \[
    a \xrightarrow{\tilde f}
    x.b \xrightarrow{x.\tilde g}
    x.y.c \xrightarrow{m_{x,y,c}}
    (x\tensor y).c\,,
    \]
  which is a \Mellies morphism from $a$ to $c$.
  \label{DefMelliesComposition}
\end{definition}
\begin{definition}[\cite{ThesisForDays}]
  Let $\X$ be a monoidal category and let 
  \[
    F,G\from \X \to \Set
    \]
  be functors.  
  Then the \emph{Day convolution} of $F$ and $G$ is a functor
  \[
    F \tensor_{\text{Day}} G \from \X \to \Set
    \]
  given by
  \[
    (F \tensor_{\text{Day}} G)(x) = \int^{y,z\from \X} F(y) \times G(z) \times \X(y\tensor z,x)\,.
    \]
  This makes $[\X,\Set]$ into a monoidal category (the monoidal unit is the functor $\X(I,\blank)\from \X \to \Set$).
  \label{DefDayConvolution}
\end{definition}

\begin{definition}[\cite{MelliesCategory}]
  Given a lax action of a monoidal category $\X$ upon a category $\C$, the \emph{\Mellies category} of the action is an $[\X,\Set]$-enriched category $\Mell_\X\C$ whose objects are the objects of $\C$ and where the hom objects are given by
  \[
    \Mell_\X\C(a,b)(x) = \C(a,x.b)\,.
    \]
  The composition is given by the natural transformation
  \begin{IEEEeqnarray*}{Cl}
    & (\Mell_\X\C(a,b) \tensor_{\text{Day}} \Mell_\X\C(b,c))(x)\\
    = & \int^{y,z\from \X} \C(a,y.b) \times \C(b,z.c) \times \X(y\tensor z,x) \\
    \to & \int^{y,z\from \X} \C(a,(y\tensor z).c) \times \X(y\tensor z,x) \\
    \cong & \C(a,x.c) \\
    = & \Mell_\X\C(a,c)(x)\,,
  \end{IEEEeqnarray*}
  where the arrow is induced by the \Mellies composition $\C(a,y.b) \times \C(b,z.c) \to \C(a,(y\tensor z).c)$.

@@

We will mainly be concerned not with the \Mellies category itself, but with a closely related ordinary category.  
The usual method for turning a $\V$-enriched category into an ordinary category is via base change along a monoidal functor $\V \to \Set$: for example, any $\V$-enriched category has an \emph{underlying ordinary category}, obtained via base change along the functor $\V(I, \blank)$.  
For example, if $a$ and $b$ are objects of the base category $\C$, then the set of morphisms in the underlying ordinary category of the \Mellies category is the set of natural transformations
\[
  \C(I,x) \to \C(a,x.b)\,,
  \]
which by the Yoneda lemma is the same as the set $\C(a,I.b)$.  
In other words, this underlying ordinary category is precisely the Kleisli category for the monad on $\C$ given by the composite
\[
  1 \xrightarrow{I}
  \X \xrightarrow{\blank.\blank}
  \End[\C]\,;
  \]
in other words, the monad on $\C$ given by $Ma = I.a$.

This is not very useful for us, since it ignores most of the action and does not give us any new theory beyond the basic theory of monads and Kleisli categories.  
We want a monoidal functor $[\M,\Set]\to\M$ that will preserve more of the structure.

Let us assume for now that $\X$ is a small category.  
Then we have a functor
\begin{IEEEeqnarray*}{cCc}
  [\X,\Set] & \to & \Set \\
  F & \mapsto & \colim_{x\from \X} F(x)
\end{IEEEeqnarray*}
given by the colimit in $\Set$.

Moreover, this functor is lax monoidal, via the morphisms
\begin{IEEEeqnarray*}{Cl}
  & \left(\int^y F(y)\right)\times \left(\int^z G(z)\right) \\
  \to & \int^{y,z} F(y) \times G(z) \\
  \to & \int^{y,z} F(y) \times G(z) \times \X(y \tensor z,x) \\
  = & \int^x (F\tensor_{\text{Day}} G)(x)
\end{IEEEeqnarray*}
(where the second morphism sends picks out the inhabitant $\inj_{y\tensor z}(\id_{y\tensor z})$ of $\colim_{x\from \X}\X(y\tensor z,x)$)
and
\begin{IEEEeqnarray*}{cCc}
  1 & \to & \colim_{x\from \X} \X(I,x) \\
  () & \mapsto & \inj_I(\id_I)\,.
\end{IEEEeqnarray*}

Rather than pursue this line of argument to its conclusion, though, we choose to define the base-changed category directly.

\begin{definition}
  Let a monoidal category $\X$ act on a category $\C$ via a lax action.  
  Then we define a new category $\C/\X$ where
  \begin{itemize}
    \item the objects are the objects of $\C$,
    \item given objects $a,b$ of $\C$, the set of morphisms $a \to b$ is given by the colimit
      \[
        \colim_{x\from \X} \C(a,x.b)\,;
        \]
      i.e., a morphism in $\C/\X$ from $a$ to $b$ is an equivalence class of \Mellies morphisms
      \[
        a \to x.b
        \]
      in $\C$, where $x$ ranges over the objects of $\X$, and where the equivalence relation on morphisms is generated by identifying two morphisms $f\from a \to x.b$, $g \from a \to y.b$ if there is a morphism $h\from x\to y$ in $\X$ such that the following diagram commutes (we say that $h$ \emph{mediates} between $f$ and $g$);
      \[
        \begin{tikzcd}
          a \arrow[r, "f"] \arrow[dr, "g"']
            & x.b \arrow[d, "h.b"] \\
          %
            & y.b
        \end{tikzcd}
        \]
    \item composition of morphisms is via the \Mellies composition of Definition \ref{DefMelliesComposition} and
    \item the identity $a\to a$ is given by the equivalence class corresponding to the morphism
      \[
        e_a \from a \to I.a
        \]
      in $\C$.
  \end{itemize}
  \label{DefCX}
\end{definition}

@@

\begin{proposition}
  $\C/\X$ is a well defined category.
\end{proposition}
\begin{proof}
  Let us first check that \Mellies composition is well defined with respect to our equivalence relation.  
  Suppose that
  \begin{gather*}
    f \from a \to x.b \\
    f' \from a \to x'.b \\
    g \from b \to y.c \\
    g' \from b \to y'.c
  \end{gather*}
  are \Mellies morphisms, and that there are morphisms $h \from x \to x'$, $k \from y \to y'$ in $\X$ such that the following diagrams commute.
  \begin{mathpar}
    \begin{tikzcd}
      a \arrow[r, "f"] \arrow[dr, "f'"']
        & x.b \arrow[d, "h.b"] \\
      %
        & x'.b
    \end{tikzcd}
    \and
    \begin{tikzcd}
      b \arrow[r, "g"] \arrow[dr, "g'"']
        & y.c \arrow[d, "k.c"] \\
      %
        & y'.c
    \end{tikzcd}
  \end{mathpar}
  Then we have the following commutative diagram, since $m$ is a natural transformation.
  \[
    \begin{tikzcd}
      a \arrow[r, "f"] \arrow[dr, "f'"']
        & x.b \arrow[r, "x.g"] \arrow[dr, "h.g" description] \arrow[d, "h.b" description]
          & x.y.c \arrow[r, "{m_{x,y,c}}"] \arrow[d, "h.y.c" description] \arrow[dd, bend left=50, "h.k.c"]
            & (x \tensor y).c \arrow[dd, "(h\tensor k).c"] \\
      %
        & x'.b \arrow[r, "x'.g" description] \arrow[dr, "x'.g'"']
          & x'.y.c \arrow[d, "x'.k.c" description]
            & \\
      %
        &
          & x'.y'.c \arrow[r, "{m_{x',y',c}}"]
            & (x' \tensor y').c
    \end{tikzcd}
    \]
  This proves that the \Mellies composition of $f$ and $g$ is equivalent to the \Mellies composition of $f'$ and $g'$.

  We should check that \Mellies composition is associative with respect to our equivalence relation.  
  Let $f \from a \to x.b$, $g\from b \to y.c$, $h \from c \to z.d$ be \Mellies morphisms.  
  Then we have the following commutative diagram.
  \[
    \begin{tikzcd}
      a \arrow[d, "f"' yshift=3pt] \arrow[d, dashed, thick]
        &
          &
            & \\
      x.b \arrow[d, "x.g"' yshift=3pt] \arrow[d, dashed, thick]
        &
          &
            & \\
      x.y.c \arrow[r, "x.y.h", thick, dashed] \arrow[d, "{m_{x,y,c}}"']
        & x.y.z.d \arrow[r, "x.m_{y,z,d}", thick, dashed] \arrow[d, "m_{x,y,z.d}"', dotted]
          & x.(y\tensor z).d \arrow[r, "m_{x,y\tensor z,d}", thick, dashed]
            & (x \tensor (y \tensor z)) . a \\
      (x \tensor y).c \arrow[r, "(x\tensor y).h"]
        & (x\tensor y).z.d \arrow[r, "m_{x\tensor y,z,a}"]
          & ((x\tensor y)\tensor z) . a \arrow[ur, "\assoc_{x,y,z}.d"', dotted]
            &
    \end{tikzcd}
    \]
  Here, the pentagon at the right is one of our conditions for a lax action, while the left-hand square commutes because $m$ is a natural transformation.  
  The composite given by the thick dashed lines is the \Mellies composition $f;(g;h)$, while that given by thin lines is the \Mellies composition $(f;g);h$.  
  The arrow $\assoc_{x,y,z}$ in $\X$ then mediates between these morphisms, so they are equivalent.

@@

\begin{remark}
  This generalizes the Kleisli category of a monad.  
  Indeed, if $M$ is a monad on $\C$, then we may consider $M$ as a lax action of the unit category on $\C$.
  Then composition of morphisms $f \from a \to I.b$, $g \from b \to I.c$ in the category $\C/1$ is given by
  \[
    a \xrightarrow{f}
    I.b \xrightarrow{I.g}
    I.I.c \xrightarrow{m_{I,I,c}}
    I.c\,.
    \]
  Identifying $I.\blank$ with $M$, we see that this is precisely the definition of composition in the Kleisli category for $M$ (and the identity is the same thing too).  
  Since the unit category has only an identity morphism, the equivalence relation on morphisms in $\C/1$ is discrete, which is why we do not need to mention it in this context.
\end{remark}

There is an identity-on-objects functor $J \from \C \to \C/\X$ given by sending a morphism $f \from a \to b$ to the \Mellies morphism
\[
  a \xrightarrow{f}
  b \xrightarrow{e_b}
  I.b\,.
  \]
It is easy to show that this is a functor, and we will prove this in a more general context in Proposition \ref{PropGrotLaxLimit}.

A special role will be played by the identity morphisms
\[
  \id_{x.a} \from x.a \to x.a\,,
  \]
considered as \Mellies morphisms $x.a \to a$.  
We will see later that these are the components of a natural transformation $\phi_{x,a}\from x.a \to a$ in $\C/\X$, generalizing the natural transformation $\phi_a \from Ma \to a$ in the Kleisli category for a monad $M$ (see Proposition \ref{pKleisli}).

We will be using the following `factorization result'.  
If $f \from a \to x.b$ is a morphism in $\C$ then the corresponding \Mellies morphism $a \to b$ may be written as the following composite in $\C/\X$.
\[
  a \xrightarrow{Jf}
  x.b \xrightarrow{\phi_{x,b}}
  b\,.
  \]
Indeed, if we compute this composite in $\C$, then we get
\[
  a \xrightarrow{f}
  x.b \xrightarrow{e_{x.b}}
  I.x.b \xrightarrow{\id}
  I.x.b \xrightarrow{m_{I,x,b}}
  (I\tensor x).b\,,
  \]
which is equivalent to $f$ in $\C/\X$, because $e_{x.b};m_{I,x,b}=\lunit_x.b$, so $\lunit_x$ mediates between this morphism and $f$.

@@

In the previous chapter, we approached the category $\C/\X$ via the \Mellies category.  
In this chapter, we will show that $\C/\X$ has important properties in its own right: namely that it is a certain lax $2$-colimit in $\Cat$.  
We will briefly return to the \Mellies category in chapter \ref{ChapPromonads}, when we will show why it is a natural idea to consider an $[\X,\Set]$-enriched category.

\begin{definition}[\cite{StreetTwoConstructions}]
  Let $\C,\D$ be bicategories.  
  A \emph{lax functor} $F\from\C \to \D$ is given by
  \begin{itemize}
    \item a map $F$ from the objects of $\C$ to the objects of $\D$;
    \item for each pair $a,b$ of objects of $\C$, a functor
      \[
        F \from \C(a,b) \to \D(F(a),F(b))\,;
        \]
    \item for each triple $a,b,c$ of objects of $\C$, a transformation 
      \[
        m_{f,g}\from F(f);F(g) \to F(f;g)
        \]
      natural in $f\from a \to b$, $g\from b \to c$; and
    \item for each object $a$ of $\C$, a $2$-cell 
      \[
        e_a\from \id_{Fa} \Rightarrow F(\id_a)\from Fa \to Fa\,;
        \]
  \end{itemize}
  such that for all tuples $a,b,c,d$ of objects and all morphisms $f\from a \to b$, $g\from b \to c$, $h \from c \to d$, the following diagrams commute.
  \begin{mathpar}
    \begin{tikzcd}[column sep=60pt]
      (F(f);F(g));F(h) \arrow[r, "{\assoc_{F(f),F(g),F(h)}}"] \arrow[d, "{m_{f,g};F(h)}"']
        & F(f);(F(g);F(h)) \arrow[d, "{F(f);m_{g,h}}"] \\
      F(f;g);F(h) \arrow[d, "{m_{f;g,h}}"']
        & F(f);F(g;h) \arrow[d, "{m_{f,g;h}}"] \\
      F((f;g);h) \arrow[r, "{F(\assoc_{f,g,h})}"]
        & F(f;(g;h))
    \end{tikzcd}
    \and
    \begin{tikzcd}
      Ff \arrow[r, "\lunit_{Ff}"] \arrow[d, "F\lunit_f"']
        & \id_{Fa} ; Ff \arrow[d, "e_a;Ff"] \\
      F(\id_a;f)
        & F\id_a;Ff \arrow[l, "{m_{\id_a,f}}"']
    \end{tikzcd}
    \and
    \begin{tikzcd}
      Ff \arrow[r, "\runit_{Ff}"] \arrow[d, "F\runit_f"']
        & Ff ; \id_{Fb} \arrow[d, "Ff;e_b"] \\
      F(f;\id_b)
        & Ff;F\id_b \arrow[l, "{m_{f,\id_b}}"']
    \end{tikzcd}
  \end{mathpar}
\end{definition}
\begin{example}
  If $\C = \BB \X$, $\D = \BB \YY$ are the delooping bicategories of monoidal categories $\X, \YY$ -- i.e., $\X$ and $\YY$ considered as bicategories with a single object -- then a lax functor $\C \to \D$ is the same thing as a lax monoidal functor $\X \to \YY$.

  More generally, a lax functor $F\from \BB \X \to \D$ is the same as a lax monoidal functor from $\X$ to the monoidal category of $1$-cells $F(*) \to F(*)$, where $*$ is the unique object of $\BB\X$.  
  In particular, a lax functor $1\to \Cat$ is the same thing as a monad and a lax functor $\BB\X \to \Cat$ is the same thing as a parametric monad parameterized by $\X$.
\end{example}

@@

\begin{remark}
  The definition of a lax colimit is normally weaker than the one we have given, in that the equations $k_c;\hat l = l_c$ and $\nu_h\hat l = \mu_h$ are not required to hold exactly, but instead up to coherent isomorphism.  
  For the lax functors we are constructing, however, we will be able to construct lax colimits that make these equations hold exactly.
\end{remark}

@@

>>>>

Equivalently, a lax action is given by a functor $\blank.\blank \from \X \times \C \to \C$ together with natural transformations
\begin{mathpar}
  m_{x,y,a}\from x.y.a \to (x\tensor y).a
  \and
  e_a \from a \to I.a
\end{mathpar}
such that the following diagrams commute.
\begin{mathpar}
  \begin{tikzcd}[column sep=36pt]
    x.y.z.a \arrow[r, "x.m_{y,z,a}"] \arrow[d, "m_{x,y,z.a}"']
      & x.(y\tensor z).a \arrow[r, "m_{x,y\tensor z,a}"]
        & (x \tensor (y \tensor z)) . a \\
    (x\tensor y).z.a \arrow[r, "m_{x\tensor y,z,a}"]
      & ((x\tensor y)\tensor z) . a \arrow[ur, "\assoc_{x,y,z}.a"']
        &
  \end{tikzcd}
  \and
  \begin{tikzcd}
    x.a \arrow[r, "e_{x.a}"] \arrow[dr, "\lunit_x.a"']
      & I.x.a \arrow[d, "m_{I,x,a}"] \\
    %
      & (I\tensor x).a
  \end{tikzcd}
  \and
  \begin{tikzcd}
    x.a \arrow[r, "x.e_a"] \arrow[dr, "\runit_x.a"']
      & x.I.a \arrow[d, "m_{x,I,a}"] \\
    %
      & (x\tensor I) . a
  \end{tikzcd}
\end{mathpar}
Since lax actions generalize monads, we shall sometimes follow \Mellies \cite{ParametricMonads} and refer to them as \emph{parametric monads}.
We shall not be considering strong actions (for which $m,e$ are required to be isomorphisms), so we shall sometimes say `action' to mean `lax action'.

@@

\begin{definition}
  Given \Mellies morphisms $f$ from $a$ to $b$ and $g$ from $b$ to $c$ given by
  \begin{mathpar}
    \tilde{f} \from a \to x.b
    \and
    \tilde{g} \from b \to y.c\,,
  \end{mathpar}
  their \emph{\Mellies composition} is given by the composite
  \[
    a \xrightarrow{\tilde f}
    x.b \xrightarrow{x.\tilde g}
    x.y.c \xrightarrow{m_{x,y,c}}
    (x\tensor y).c\,,
    \]
  which is a \Mellies morphism from $a$ to $c$.
  \label{DefMelliesComposition}
\end{definition}
\begin{definition}[\cite{ThesisForDays}]
  Let $\X$ be a monoidal category and let 
  \[
    F,G\from \X \to \Set
    \]
  be functors.  
  Then the \emph{Day convolution} of $F$ and $G$ is a functor
  \[
    F \tensor_{\text{Day}} G \from \X \to \Set
    \]
  given by the coend
  \[
    (F \tensor_{\text{Day}} G)(x) = \int^{y,z\from \X} F(y) \times G(z) \times \X(y\tensor z,x)\,.
    \]
  This makes $[\X,\Set]$ into a monoidal category (the monoidal unit is the functor $\X(I,\blank)\from \X \to \Set$).
  \label{DefDayConvolution}
\end{definition}

\begin{definition}[\cite{MelliesCategory}]
  Given a lax action of a monoidal category $\X$ upon a category $\C$, the \emph{\Mellies category} of the action is an $[\X,\Set]$-enriched category $\Mell_\X\C$ whose objects are the objects of $\C$ and where the hom objects are given by
  \[
    \Mell_\X\C(a,b)(x) = \C(a,x.b)\,.
    \]
  The composition is given by the natural transformation
  \begin{IEEEeqnarray*}{Cl}
    & (\Mell_\X\C(a,b) \tensor_{\text{Day}} \Mell_\X\C(b,c))(x)\\
    = & \int^{y,z\from \X} \C(a,y.b) \times \C(b,z.c) \times \X(y\tensor z,x) \\
    \to & \int^{y,z\from \X} \C(a,(y\tensor z).c) \times \X(y\tensor z,x) \\
    \cong & \C(a,x.c) \\
    = & \Mell_\X\C(a,c)(x)\,,
  \end{IEEEeqnarray*}
  where the arrow on the third line is induced by the \Mellies composition $\C(a,y.b) \times \C(b,z.c) \to \C(a,(y\tensor z).c)$.

@@

We will mainly be concerned not with the \Mellies category itself, but with a closely related ordinary category.  
The usual method for turning a $\V$-enriched category into an ordinary category is via base change along a monoidal functor $\V \to \Set$.

For example, any $\V$-enriched category has an \emph{underlying ordinary category}, obtained via base change along the functor $\V(I, \blank)$.  
If $a$ and $b$ are objects of the base category $\C$, then the set of morphisms in the underlying ordinary category of the \Mellies category is the set of natural transformations
\[
  \C(I,x) \to \C(a,x.b)\,,
  \]
which by the Yoneda lemma is the same as the set $\C(a,I.b)$.  
In other words, this underlying ordinary category is precisely the Kleisli category for the monad on $\C$ given by the composite
\[
  1 \xrightarrow{I}
  \X \xrightarrow{\blank.\blank}
  \End[\C]\,;
  \]
in other words, the monad on $\C$ given by $Ma = I.a$.

More generally, if $m\tensor m\to m$ is a monoid in $\X$, then we can construct a lax monoidal functor $[\X,\Set] \to \Set$ that sends $F$ to $F(m)$.  
If we change base through this functor, then the morphisms $a \to b$ in the resulting ordinary category will be elements of $\C(a,m.b)$; i.e., morphisms $a \to m.b$ in $\C$, and the category turns out to be the Kleisli category corresponding to the monad on $\C$ given by $Ma = m.a$.
So the \Mellies category can be thought of as classifying all the Kleisli categories that arise from the action $\blank.\blank$ through the monoids in $\M$.

For our purposes, however, we will want to use a single ordinary category that captures as much as possible of the structure of the \Mellies category.
Let us assume for now that $\X$ is a small category.  
Then we have a functor
\begin{IEEEeqnarray*}{cCc}
  [\X,\Set] & \to & \Set \\
  F & \mapsto & \colim_{x\from \X} F(x)
\end{IEEEeqnarray*}
given by the colimit in $\Set$.

Moreover, this functor is lax monoidal, via the multiplicative
\begin{IEEEeqnarray*}{Cl}
  & \left(\int^y F(y)\right)\times \left(\int^z G(z)\right) \\
  \to & \int^{y,z} F(y) \times G(z) \\
  \to & \int^{y,z} F(y) \times G(z) \times \X(y \tensor z,x) \\
  = & \int^x (F\tensor_{\text{Day}} G)(x)
\end{IEEEeqnarray*}
(where the second morphism sends picks out the inhabitant $\inj_{y\tensor z}(\id_{y\tensor z})$ of $\colim_{x\from \X}\X(y\tensor z,x)$),
and via the unital
\begin{IEEEeqnarray*}{cCc}
  1 & \to & \colim_{x\from \X} \X(I,x) \\
  () & \mapsto & \inj_I(\id_I)\,.
\end{IEEEeqnarray*}

We then conclude by changing base through this colimit functor.  
It turns out to be fairly easy to describe the resulting category directly.

\begin{definition}
  Let a monoidal category $\X$ act on a category $\C$ via a lax action.  
  Then we define a new category $\C/\X$ where
  \begin{itemize}
    \item the objects are the objects of $\C$;
    \item given objects $a,b$ of $\C$, the set of morphisms $a \to b$ is given by the colimit
      \[
        \colim_{x\from \X} \C(a,x.b)\,;
        \]
      i.e., a morphism in $\C/\X$ from $a$ to $b$ is an equivalence class of \Mellies morphisms
      \[
        a \to x.b
        \]
      in $\C$, where $x$ ranges over the objects of $\X$, and where the equivalence relation on morphisms is generated by identifying two morphisms $f\from a \to x.b$, $g \from a \to y.b$ if there is a morphism $h\from x\to y$ in $\X$ such that the following diagram commutes (we say that $h$ \emph{mediates} between $f$ and $g$);
      \[
        \begin{tikzcd}
          a \arrow[r, "f"] \arrow[dr, "g"']
            & x.b \arrow[d, "h.b"] \\
          %
            & y.b
        \end{tikzcd}
        \]
    \item composition of morphisms is via the \Mellies composition of Definition \ref{DefMelliesComposition}; and
    \item the identity $a\to a$ is given by the equivalence class corresponding to the morphism
      \[
        e_a \from a \to I.a
        \]
      in $\C$.
  \end{itemize}
  \label{DefCX}
\end{definition}

@@

\begin{proposition}
  $\C/\X$ is a well defined category.
\end{proposition}
\begin{proof}
  Let us first check that \Mellies composition is well defined with respect to our equivalence relation.  
  Suppose that
  \begin{gather*}
    f \from a \to x.b \\
    f' \from a \to x'.b \\
    g \from b \to y.c \\
    g' \from b \to y'.c
  \end{gather*}
  are \Mellies morphisms, and that $f$ is equivalent to $f'$ and $g$ is equivalent to $g'$.  
  By induction, we can assume that $f,f'$ and $g,g'$ are related by the relation that generates the equivalence relation on \Mellies morphisms, so that there are morphisms $h \from x \to x'$, $k \from y \to y'$ in $\X$ such that the following diagrams commute.
  \begin{mathpar}
    \begin{tikzcd}
      a \arrow[r, "f"] \arrow[dr, "f'"']
        & x.b \arrow[d, "h.b"] \\
      %
        & x'.b
    \end{tikzcd}
    \and
    \begin{tikzcd}
      b \arrow[r, "g"] \arrow[dr, "g'"']
        & y.c \arrow[d, "k.c"] \\
      %
        & y'.c
    \end{tikzcd}
  \end{mathpar}
  Then we have the following commutative diagram, since $m$ is a natural transformation.
  \[
    \begin{tikzcd}
      a \arrow[r, "f"] \arrow[dr, "f'"']
        & x.b \arrow[r, "x.g"] \arrow[dr, "h.g" description] \arrow[d, "h.b" description]
          & x.y.c \arrow[r, "{m_{x,y,c}}"] \arrow[d, "h.y.c" description] \arrow[dd, bend left=50, "h.k.c"]
            & (x \tensor y).c \arrow[dd, "(h\tensor k).c"] \\
      %
        & x'.b \arrow[r, "x'.g" description] \arrow[dr, "x'.g'"']
          & x'.y.c \arrow[d, "x'.k.c" description]
            & \\
      %
        &
          & x'.y'.c \arrow[r, "{m_{x',y',c}}"]
            & (x' \tensor y').c
    \end{tikzcd}
    \]
  This proves that the \Mellies composition of $f$ and $g$ is equivalent to the \Mellies composition of $f'$ and $g'$.

  We should check that \Mellies composition is associative with respect to our equivalence relation.  
  Let $f \from a \to x.b$, $g\from b \to y.c$, $h \from c \to z.d$ be \Mellies morphisms.  
  Then we have the following commutative diagram.
  \[
    \begin{tikzcd}
      a \arrow[d, "f"' yshift=3pt] \arrow[d, dashed, thick]
        &
          &
            & \\
      x.b \arrow[d, "x.g"' yshift=3pt] \arrow[d, dashed, thick]
        &
          &
            & \\
      x.y.c \arrow[r, "x.y.h", thick, dashed] \arrow[d, "{m_{x,y,c}}"']
        & x.y.z.d \arrow[r, "x.m_{y,z,d}", thick, dashed] \arrow[d, "m_{x,y,z.d}"', dotted]
          & x.(y\tensor z).d \arrow[r, "m_{x,y\tensor z,d}", thick, dashed]
            & (x \tensor (y \tensor z)) . a \\
      (x \tensor y).c \arrow[r, "(x\tensor y).h"]
        & (x\tensor y).z.d \arrow[r, "m_{x\tensor y,z,a}"]
          & ((x\tensor y)\tensor z) . a \arrow[ur, "\assoc_{x,y,z}.d"', dotted]
            &
    \end{tikzcd}
    \]
  Here, the pentagon at the right is one of the coherence diagrams for a lax action, while the left-hand square commutes because $m$ is a natural transformation.  
  The composite given by the thick dashed lines is the \Mellies composition $f;(g;h)$, while that given by thin lines is the \Mellies composition $(f;g);h$.  
  The arrow $\assoc_{x,y,z}$ in $\X$ then mediates between these morphisms, so they are equivalent, and therefore correspond to the same morphism in $\C/\X$.

@@

\begin{remark}
  This generalizes the Kleisli category of a monad.  
  Indeed, if $M$ is a monad on $\C$, then we may consider $M$ as a lax action of the unit category $1$ on $\C$.
  Then composition of morphisms $f \from a \to I.b$, $g \from b \to I.c$ in the category $\C/1$ is given by
  \[
    a \xrightarrow{f}
    I.b \xrightarrow{I.g}
    I.I.c \xrightarrow{m_{I,I,c}}
    I.c\,.
    \]
  Writing $M$ for $I.\blank$, we see that this is precisely the definition of composition in the Kleisli category for $M$ (and the identity is the same thing too).  
  Note that since the unit category has only an identity morphism, the equivalence relation on morphisms in $\C/1$ is discrete, so $\C/1$ \emph{is} the Kleisli category for $M$.
\end{remark}

There is an identity-on-objects functor $J \from \C \to \C/\X$ given by sending a morphism $f \from a \to b$ to the \Mellies morphism
\[
  a \xrightarrow{f}
  b \xrightarrow{e_b}
  I.b\,.
  \]
It is easy to show that this is a functor, and we will prove that it is in a more general context in Proposition \ref{PropGrotLaxLimit}.

A special role will be played by the identity morphisms
\[
  \id_{x.a} \from x.a \to x.a\,,
  \]
considered as \Mellies morphisms $x.a \to a$.  
We will see later that these are the components of a natural transformation $\phi_{x,a}\from x.a \to a$ in $\C/\X$, generalizing the natural transformation $\phi_a \from Ma \to a$ in the Kleisli category for a monad $M$ that we met in Proposition \ref{pKleisli}.

We will be meeting the following `factorization result', analogous to the corresponding result for Kleisli categories that we met in the previous section.
If $f \from a \to x.b$ is a morphism in $\C$ then the corresponding \Mellies morphism $a \to b$ may be written as the following composite in $\C/\X$.
\[
  a \xrightarrow{Jf}
  x.b \xrightarrow{\phi_{x,b}}
  b\,.
  \]
Indeed, if we compute this composite in $\C$, then we get
\[
  a \xrightarrow{f}
  x.b \xrightarrow{e_{x.b}}
  I.x.b \xrightarrow{\id}
  I.x.b \xrightarrow{m_{I,x,b}}
  (I\tensor x).b\,,
  \]
which is equivalent to $f$ in $\C/\X$, because $e_{x.b};m_{I,x,b}=\lunit_x.b$, so $\lunit_x$ mediates between this morphism and $f$.

@@

In the previous chapter, we approached the category $\C/\X$ via the \Mellies category.  
In this chapter, we will show that $\C/\X$ has important properties in its own right: namely that it is a certain lax $2$-colimit in $\Cat$.  
We will briefly return to the \Mellies category in chapter \ref{ChapPromonads}, when we will approach it from the point of view of profunctors.

\begin{definition}[\cite{StreetTwoConstructions}]
  Let $\C,\D$ be bicategories.  
  A \emph{lax functor} $F\from\C \to \D$ is given by
  \begin{itemize}
    \item a map $F$ from the objects of $\C$ to the objects of $\D$;
    \item for each pair $a,b$ of objects of $\C$, a functor
      \[
        F \from \C(a,b) \to \D(F(a),F(b))\,;
        \]
    \item for each triple $a,b,c$ of objects of $\C$, a transformation 
      \[
        m_{f,g}\from F(f);F(g) \to F(f;g)
        \]
      natural in $f\from a \to b$, $g\from b \to c$; and
    \item for each object $a$ of $\C$, a $2$-cell 
      \[
        e_a\from \id_{Fa} \Rightarrow F(\id_a)\from Fa \to Fa\,;
        \]
  \end{itemize}
  such that for all tuples $a,b,c,d$ of objects and all morphisms $f\from a \to b$, $g\from b \to c$, $h \from c \to d$, the following diagrams commute.
  \begin{mathpar}
    \begin{tikzcd}[column sep=60pt]
      (F(f);F(g));F(h) \arrow[r, "{\assoc_{F(f),F(g),F(h)}}"] \arrow[d, "{m_{f,g};F(h)}"']
        & F(f);(F(g);F(h)) \arrow[d, "{F(f);m_{g,h}}"] \\
      F(f;g);F(h) \arrow[d, "{m_{f;g,h}}"']
        & F(f);F(g;h) \arrow[d, "{m_{f,g;h}}"] \\
      F((f;g);h) \arrow[r, "{F(\assoc_{f,g,h})}"]
        & F(f;(g;h))
    \end{tikzcd}
    \and
    \begin{tikzcd}
      Ff \arrow[r, "\lunit_{Ff}"] \arrow[d, "F\lunit_f"']
        & \id_{Fa} ; Ff \arrow[d, "e_a;Ff"] \\
      F(\id_a;f)
        & F\id_a;Ff \arrow[l, "{m_{\id_a,f}}"']
    \end{tikzcd}
    \and
    \begin{tikzcd}
      Ff \arrow[r, "\runit_{Ff}"] \arrow[d, "F\runit_f"']
        & Ff ; \id_{Fb} \arrow[d, "Ff;e_b"] \\
      F(f;\id_b)
        & Ff;F\id_b \arrow[l, "{m_{f,\id_b}}"']
    \end{tikzcd}
  \end{mathpar}
\end{definition}
\begin{example}
  If $\C = \BB \X$, $\D = \BB \YY$ are the delooping bicategories of monoidal categories $\X, \YY$ -- i.e., $\X$ and $\YY$ considered as bicategories with a single object -- then a lax functor $\C \to \D$ is the same thing as a lax monoidal functor $\X \to \YY$.  
\end{example}

\begin{example}
  More generally, a lax functor $F\from \BB \X \to \D$ is the same as a lax monoidal functor from $\X$ to the monoidal category of $1$-cells $F(*) \to F(*)$, where $*$ is the unique object of $\BB\X$.  
  In particular, a lax functor $1\to \Cat$ is the same thing as a monad and a lax functor $\BB\X \to \Cat$ is the same thing as a parametric monad parameterized by $\X$.
\end{example}

@@

\begin{remark}
  The definition of a lax colimit is normally weaker than the one we have given, in that the equations $k_c;\hat l = l_c$ and $\nu_h\hat l = \mu_h$ are not required to hold exactly, but instead up to coherent isomorphism.  
  The distinction is not important for our purposes, since the lax colimits we will be considering will all satisfy our stronger definition.
\end{remark}

We shall sometimes write
\[
  \laxcolim_{c \from \C}F(c)
  \]
for the lax colimit of the functor $F$.

@@

But first, an interlude.
